Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-01-21 article/article.Rmd
* reduce disk usage of analysis * progress writing manuscript

---
title: "Are environmental and geographic effective surrogates for genetic variation in conservation planning?"
author: |
  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
  | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
  | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: ["conservation", "biodiversity", "AFLP", "reserve-selection"]
abstract: "Insert abstract here."
output:
  rmarkdown::pdf_document:
    toc: false
    toc_depth: 5
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: preamble-latex.tex
      after_body: postamble-latex.tex
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
---

@@

Our overarching aim is to investigate the effectiveness of environmental and geographic surrogates for conservation planning. First, we aim to test the effectiveness of these surrogates under the simplest of conditions by using a set of single-species prioritisations. Second, we aim to test the effectiveness of these surrogates in a more realistic multi-species context, and see how they compare to the single-species prioritisations. Third, we aim to understand the relationship between the proportion of surrogate variation secured in a multi-species prioritisation and the prorpotion of genetic variation secured in a prioritization. If these surrogates markedly improve the proportion of genetic variation secured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.

@@

### Prioritizations
We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465]. This formulation can identify the minimum number of planning units required to preserve a proportion of the species' range as well as a proportion of the species' variation using an attribute space [_sensu_ @r465]. We treated the genomic and surrogate variables as distinct attribute spaces. We set the demand points for each species and its attribute space using the values associated with the planning units.

To fulfil the aims of the study, we generated three sets of prioritisations. First, we generated a set of single species prioritizations. For each species, we generated `r rapr.params.LST[[MODE]]$single.species.amount.replicates` prioritizations using a `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based target, a single prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based target and `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % space-based targets for the surrogate variation, and single a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based target and `r (rapr.params.LST[[MODE]]$genetic.target*100) ` % space-based targets for the genomic variation. We generated multiple solutions for the single-species prioritizations containing only amount-based targets because they were effectively a random selection of the planning units that the species inhabits, and so we obtained multiple solutions to account for stochasticity. Next, we generated three multi-species prioritizations using the same sets of targets used in the single species prioritizations. Finally, to understand the relationship between the surrogate-based and genomic-based targets, we generated a suite of prioritizations using increasing space-based targets between `r (min(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$pareto.surrogate.targets[1:2])*100` %. All of these prioritisations had no amount-based targets. For each prioritization we calculated the proportion of the adaptive (if present) and neutral genetic variation that it secured. We solved the reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

## Statistical analysis
We tested for differences in the effectiveness of prioritizations at securing genetic variation using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469]. We fit two full models to the proportion of genomic variation secured in the single- and multi-species prioritizations using a logit link. Both models contained a predictor variable indicating if the prioritisation was generated using amount, surrogate, or genetic targets, a second predictor variable indicating if the genomic variation was adaptive or neutral, and the interaction between these variables. Additionally, both models contained an intercept-only random effect to ensure that comparisons were made within species. Each full model was subject to a stepwise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$). Finally, we conducted a post-hoc analysis using Tukey constrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].

>>>>

---
title: "Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?"
author: |
  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
  | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
  | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: ["conservation", "biodiversity", "AFLP", "reserve-selection"]
abstract: "Insert abstract here."
output:
  rmarkdown::pdf_document:
    toc: false
    toc_depth: 5
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: preamble-latex.tex
      after_body: postamble-latex.tex
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
---

@@

Our overarching aim is to investigate the effectiveness of environmental and geographic surrogates for adaptive neutral genetic variation in the context of conservation planning. First, we aim to test the effectiveness of these surrogates under the simplest of conditions by using a set of single-species prioritisations. Second, we aim to test the effectiveness of these surrogates in a more realistic multi-species context, and see how they compare to the single-species prioritisations. Third, we aim to understand the relationship between the proportion of surrogate variation secured in a multi-species prioritisation and the prorpotion of genetic variation secured in a prioritization. If these surrogates can be shown to markedly improve the proportion of genetic variation secured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.

@@

### Prioritizations
We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465]. This formulation can identify the minimum number of planning units required to preserve a proportion of the species' range as well as a proportion of the species' variation using an attribute space [_sensu_ @r465]. We treated the genomic and surrogate variables as distinct attribute spaces. We set the demand points for each species and its attribute space using the values associated with the planning units. Although Hanson _et al._ [-@r465] recomend against using the same coordinates for demand points and planning units for real-word conservation planning exercises, we did this to provide a more robust test and reduce uncertainty in the analysis. For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

To understand how surrogate-based targets might perform under typical conservation planning scenarios, we generated a set of single-species and multi-species prioritizations. All of these prioritizations used `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based targets to ensure an adequate proportion of habitat was secured for each species [@r197; @r255; @r471]. We generated a set of single-species prioritizations for each species: `r rapr.params.LST[[MODE]]$single.species.amount.replicates` prioritizations using only the amount-based targets, a single prioritization using the amount-based target and `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % surrogate-based targets, and a single prioritization using the amount-based target and `r (rapr.params.LST[[MODE]]$genetic.target*100)` % genetic-based targets. We also generated three multi-species prioritizations. These prioritizations used the three unique combinations of targets used in the single-species prioritizations. Note that we obtained multiple solutions for the single-species prioritisations that only used an amount-based target because, unlike the prioritizations, these targets yielded many optimal solutions with varying levels effectiveness.

To understand the relationship between the surrogate-based targets and the proportion of genetic variation secured, we generated a set of multi-species prioritizations with increasing surrogate-based targets between `r (min(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$pareto.surrogate.targets[1:2])*100` %. None of these priortizations used amount-based targets.

## Statistical analysis
We tested for differences in the effectiveness of prioritizations based on typical scenarios using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469]. We fit two full models to the proportion of genomic variation secured in the single- and multi-species prioritizations using a logit link. Both models contained a predictor variable indicating if the prioritisation was generated using amount, surrogate, or genetic targets, a second predictor variable indicating if the genomic variation was adaptive or neutral, and the interaction between these variables. Additionally, both models contained an intercept-only random effect to ensure that comparisons were made within species. Each full model was subject to a stepwise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$). Finally, we conducted a post-hoc analysis using Tukey constrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].

We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models [NLMM; using the $\texttt{lme4}$ R package]. Two models were fit to the adaptive and neutral genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = 0 + (0 - 1) \times exp(r \times x)$; $r$ is the accumulate rate). Both models contained intercept and slope random effects to accomodate variation between species.
